### 2021-07-25 article/articleCustomerDropoutMembership.Rmd
upd

---
title: |
  |
  |
  |
  | \vspace{1cm}Customer dropout membership^[Corresponding address: sobreiro@esdrm.ipsantarem.pt. The current template adapts part of the Rmd code by [Paul C. Bauer](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown), Mannheim Centre for European Social Research.]\vspace{0.5cm}
author: |
  | Pedro Sobreiro, Javier Berrocal, Domingos Martinho, José Garcia Alonso
  |
  | Extremadura University
  |
date: |
  |
  |
  | `r gsub("^0", "", format(Sys.time(), "%d %B, %Y"))`
  |
  |
linestretch: 1.2
colorlinks: true
abstract: \noindent\setstretch{1}Abstract of the article. Here we can place more info.\vspace{.8cm}
bibliography: references.bib
csl: american-sociological-association.csl
output:
  bookdown::pdf_document2:
    includes:
    toc: no
    keep_tex: true
    latex_engine: xelatex
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin = 1in
always_allow_html: yes
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{color}
   - \usepackage{pdfpages}
   - \usepackage{amsmath}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{hyperref}
editor_options: 
  markdown: 
    wrap: 80
---

```{r Setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use cache = TRUE if you want to speed up compilation
# set path
# get rmarkdown directory
caminho <- getwd()
# set working directory
setwd(caminho)
print(caminho)

# A function to allow for showing some of the inline code
rinline <- function(code) {
  html <- '<code  class = "r">``` `r CODE` ```</code>'
  sub("CODE", code, html)

##https://opensource.com/article/19/5/python-3-default-mac
#Sys.setenv(RETICULATE_PYTHON = "C:/Users//AppData/Local/Microsoft/WindowsApps/python3")

#install.packages("reticulate")
#library(reticulate)
#virtualenv_create("artigosurvival",python_version="3.7.10")
#use_virtualenv("artigosurvival", required = TRUE)
}   
```

```{r RETICULATE,echo = FALSE}
#set environment first
Sys.setenv(RETICULATE_PYTHON="C:\\Users\\sobre\\anaconda3\\envs\\survival\\python.exe")
#call reticulate
library(reticulate)
#activate environment
use_condaenv("survival", required = TRUE)
#use_virtualenv("C:/Users/mangelo.EEG/Documents/python")
```

# Introduction

Research idea:

-

Context: An organization membership located in Portugal. The organization offers
an annual membership for the members, the service subscription has several
payment options:

-   Men with a annual fee of 10€
-   Women annual fee of 6€
-   Correspondent fee 6€
-   Retired fee 5€
-   Student fee 2.5€
-   under-14 fee 1€

\clearpage

\renewcommand{\baselinestretch}{0.5}\normalsize

\renewcommand{\baselinestretch}{1.1}\normalsize

\clearpage

```{r Stats1, echo = TRUE, message = FALSE}
library(dplyr)
library(dlookr)
library(ggplot2)

#eda_report(nlswork,output_dir  = 
#	"C:/Users/mangelo.EEG/Documents/GitHub/prjs/reports/",
#	output_file = "eda_report.pdf")

## The data

names(airquality)
#summary(nlswork)

## Missing values
library(visdat)
vis_dat(airquality)

library(naniar)
vis_miss(airquality)

gg_miss_upset(airquality)
## GRAPHS
dplyr::glimpse(cars$Ozone)
d <- density(airquality$Temp)
plot(d)

plot(airquality$Ozone, airquality$Solar.R)
ggplot(airquality, aes(x = Ozone, y = Solar.R)) +
geom_miss_point()

ggplot(airquality, aes(x = Wind, y = Temp)) +
geom_miss_point() +
facet_wrap(vars(Month))

stats <- summary(airquality$Temp)
stats

describe(airquality)

```

# Experimental Results

## Data description

```{r get_data, echo = FALSE, message = FALSE, warning=FALSE}
library(stargazer)
library(readxl)
library(dplyr)
library(visdat)

df_members <- read_excel("../data/membershipData.xlsx")

names(df_members)
# rename column labels
names(df_members) <- c("num_socio", "dt_inscription", "year", "birth_date",
                       "age", "sex", "marital_status", "category",
                       "monthly_fee", "occupation", "zip_code",
                       "dt_last_invoice", "dt_last_payment", "total_amount",
                       "total_matches", "season_matches",
                       "days_since_last_payment", "months_since_last_payment",
                       "dropout", "years_membership", "stadium_access",
                       "quart_stadium_entries", "inscription_month")

names(df_members)

# select relevant variables
df_members <- df_members %>%
select(year, age, sex, marital_status, monthly_fee, total_amount, total_matches,
       season_matches, months_since_last_payment, dropout, years_membership,
       stadium_access, quart_stadium_entries,inscription_month)

str(df_members)
vis_dat(df_members) #check
```

Table \@ref(tab:tab1cars) shows data's summary statistics.[^1] `stargazer()` is
and excellent solution to export outputs.

[^1]: You can reference the table as \@ref(tab1cars).

```{r Export Stats0, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(stargazer)
df_summary <- summary(df_members, object)
stargazer(df_members,
          title = "Summary statistics",
          label = "tab1cars",
          table.placement = "h",
          header = FALSE,
          summary = TRUE,
          summary.stat = c("min", "p25", "median", "p75", "max", "median", "sd")
         )

```

Teste \@ref(tab:summarytable)

```{r summarytable, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(gtsummary)
library(kableExtra)

tbl <- df_members %>%
    tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})",
                                 all_categorical() ~ "{p}%")
               )
as_kable_extra(tbl,booktabs=T,caption="Summary statistics")
```

```{python Py,eval=TRUE,echo=TRUE}
from pysurvival.utils.display import correlation_matrix
import pandas as pd
import numpy as np

col = ['sex','marital_status','quart_stadium_entries']

df_members = r.df_members #copy r dataframe to python

df_members = pd.get_dummies(df_members, columns=col,drop_first=True)

# Creating the time and event columns
time_column = 'years_membership'
event_column = 'dropout'

# Extracting the features
features = np.setdiff1d(df_members.columns, [time_column, event_column] ).tolist()

correlation_matrix(df_members, figure_size=(10,10), text_fontsize=6)
```

The average age in our data is `r round(mean(df_members$age,na.rm = TRUE),1)`.

# Tables {#sec:tables}

R Markdown PDF is now able to produce good tables with our output. For
`stargazer` the label is contained in the function, while for `kable` it's
contained in the chunk name.

## stargazer(): Summary and regression tables

Table \@ref(tab2) reports regression outputs. Name the models as you can refer
to their names in the text (M1, M2, M3).

```{r Export Regressions, echo = TRUE, message = FALSE, warning = FALSE, paged.print = FALSE, results = "asis"}
library(stargazer)
model1 <- lm(speed ~ dist, data = cars)
model2 <- lm(speed ~ dist, data = cars)
model3 <- lm(dist ~ speed, data = cars)
stargazer(model1, model2, model3,
          title = "Regression table with stargazer",
          label = "tab2",
          table.placement = "h",
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header = FALSE)
```

# Figures

## Graphs with R

You can insert figures like this. One would like to produce and insert them on
the fly in the `.rmd` file. Figure \@ref(fig:fig-1) is such an example.

```{r Figures 1, fig-1, fig.align = "center", fig.cap = "Scatterplot of Speed and Distance", fig.pos = "h", message = FALSE, warning = FALSE, paged.print = FALSE}
plot(cars$speed, cars$dist)
```

However, in some cases it does not work.

## Example: ggplot2 graphs

See the `ggplot2` output reported in Figure \@ref(fig:fig-2).

```{r Fig-2, fig.align = "center", fig.cap = "Miles per gallon according to the weight", fig.pos = "h", fig.width = 6, fig.height = 3, message = FALSE, warning = FALSE, paged.print = FALSE}
mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg, shape = cyl)) + geom_point() +
  labs(x = "Weight (lb/1000)", y = "Miles/(US) gallon",
       shape = "Number of \n Cylinders") + theme_classic()
```

## Another example using Plotly

With `Plotly` we can produce interactive graphs which play well, for example,
once can embedded in html webpages (drop by
[here](https://paulcbauer.shinyapps.io/visualizing-causal-scenarios/) for an
example). One can insert this type of graphs in R Markdown PDF using `Orca` (it
generates static images from Plotly graphs). Go
[here](https://github.com/plotly/orca#installation) to check how to install it.
See Figure \@ref(fig:fig-3) for an example.

```{r Fig-3, message = FALSE, warning = FALSE}
library(plotly)
p <- plot_ly(cars, type = "scatter", mode = "markers",
        x = ~speed,
        y = ~dist)
#Sys.setenv('MAPBOX_TOKEN' = '12423423') # set arbitrary token
#orca(p, "logs/plotly-plot.pdf")
```

```{=tex}
\begin{figure}[ht]
\centering
\caption{Example: export a Plotly figure using `orca`}\label{fig:fig-3}
    \includegraphics[width = 0.9\linewidth]{../figures/teste.png}
\begin{flushleft}
\end{flushleft}
\end{figure}
\vspace{-1.2cm}
```
```{r exemploMedia}
# Lets create a value for example

media <- mean(cars$speed)

```

The criminal rate is `r media`%o.

\vspace{0.3cm}

# Miguel's tests

## R

Example of an equation

$$\int_0^{2\pi} \sin x~dx$$

*Example of a matrix*

$$
\mathbf{X} = \left[\begin{array}
{rrr}
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9
\end{array}\right]
$$

or

```{=tex}
\begin{equation}
f\left(k\right) = \binom{n}{k}p^k\left(1-p\right)^{n-k} \label{eq:binom}
\end{equation}
```
See Equation \@ref(eq:binom).

```{=tex}
\begin{align}
y_{ijt} = \beta x_{ijt} + \eta_i + \gamma_j + \lambda_t + \varepsilon_{ijt}
\end{align}
```
```{r echo = TRUE, message = FALSE, warning = FALSE, paged.print = FALSE, results = "asis"}
library(stargazer)
stargazer(cars,
          title = "Summary 24",
          label = "tab24",
          table.placement = "ht",
          header = FALSE)
```

# Final remarks

Check the replication package for Bonhomme, Lamadon and Manresa (2019):
<https://github.com/tlamadon/blm-replicate>

\newpage

# References {.unnumbered}

::: {#refs}
:::

# Appendix: Chunk options {.unnumbered}

## Software versioning

### R

```{r fig-versioning, echo = TRUE}
cat(paste("#", capture.output(sessionInfo()), "\n", collapse  = ""))
  # or use message() instead of cat()
```