### 2024-01-11 article/customer_fitness_artigo.Rmd
update

---
title: |
  |
  |
  |
  | \vspace{1cm}Health club customer dropout membership^[Corresponding address: sobreiro@esdrm.ipsantarem.pt. Quality of Life Research Centre, Polytechnic Institute of Santarém, Portugal]\vspace{0.5cm}
author: |
  | Pedro Sobreiro, Javier Berrocal, Domingos Martinho, José Garcia Alonso
  |
  | Extremadura University
  |
date: |
  |
  |
  | `r gsub("^0", "", format(Sys.time(), "%d %B, %Y"))`
  |
  |
linestretch: 1.2
colorlinks: true
abstract: \vspace{1cm} Customer retention is fundamental to improve the organizations profits. Existing approaches normaly explore static models predicting when the customer will dropout. In this paper we propose a different approach considering that the customer dropout risk changes over time using clusters to improve the model performance. We explore a survival model using random forests with and with out clusters in a dataset of 5209 customers. The model using clusters improved the performance significantly. This paper shows that the use of clusters should be considered to identify dropout patterns to support the timming when the dropout occurs considering the cluster where the customer is. Improving the performance as less errors and is expected to contribute to the identification when should be developed retention strategies. \vspace{.8cm} 
bibliography: references.bib
csl: american-sociological-association.csl
output:
  bookdown::pdf_document2:
    includes:
    toc: no
    keep_tex: true
    latex_engine: xelatex
    number_sections: true
  bookdown::word_document2:
    latex_engine: xelatex
    number_sections: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin = 1in
always_allow_html: yes
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{color}
   - \usepackage{pdfpages}
   - \usepackage{amsmath}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{hyperref}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
editor_options: 
  markdown: 
    wrap: 80
---

```{r Setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE, warning = FALSE, message = FALSE)
# Use cache = TRUE if you want to speed up compilation
# set path
# get rmarkdown directory
caminho <- getwd()
# set working directory
setwd(caminho)
print(caminho)
# A function to allow for showing some of the inline code
rinline <- function(code) {
  html <- '<code  class = "r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

```{r RETICULATE, message=FALSE, warning=FALSE, include=FALSE}
# load essential libraries
library(dplyr)
library(dlookr)
library(ggplot2)
library(reticulate)
#Replace by your environment, usually which python solves the problem
#conda env list also is a good option
#set environment first then call reticulate library
path_python_windows <- "C:\\Users\\sobre\\AppData\\Local\\r-miniconda\\envs\\rsurvival\\python.exe"
path_python_linux <- "/home/sobreiro/miniconda3/envs/survival/bin/python"
switch(Sys.info()[["sysname"]],
       Windows = {
            Sys.setenv(RETICULATE_PYTHON = path_python_windows)
            #call reticulate
            library(reticulate)
            #activate environment
            use_condaenv("rsurvival", required = TRUE)
            caminho_figuras <- "c:/nuvem/Dropbox/doutoramento/tese/9.tese_documento/PhD-Pedro-Sobreiro/figures/"
       },
       Linux = {
            Sys.setenv(RETICULATE_PYTHON = path_python_linux)
            library(reticulate)
            use_condaenv("survival", required = TRUE)
            matplotlib <- import("matplotlib")
            matplotlib$use("Agg", force = TRUE)
            # https://community.rstudio.com/t/reticulate-rstudio-matplotlib/33219
            caminho_figuras <- "/mnt/c/nuvem/Dropbox/doutoramento/tese/9.tese_documento/PhD-Pedro-Sobreiro/figures/"
       }
)
```

# Introduction

Customer retention is a problem is a problem being addressed using the dropout 
prediction as an insight to identify customers that could dropout.
The customers database is the most valuable asset that the organizations possess
[@Athanassopoulos_2000].
The problem of retention is not new, in a seminal paper @Copeland_1923 address
the problem of brand loyalty and in marketing research [@Mellens_Dekimpe_Steenkamp_1996].

The advantage of developing some strategies in retention are supported in the idea
that the costs of customer retention are lower than customer acquisition 
[@Fornell_Wernerfelt_1987; @Edward_Sahadev_2011].
The increase in the profits with the reduction of 5\% of the dropout could represent
almost a duplication of the profits [@Reichheld_1996].

The development of a customer retention strategy could be supported in the 
identification of the customers that will dropout [@alboukaey_dynamic_2020]. 
However, dropout has two underlying scenarios contractual and non-contractual 
settings {@Gupta2006Modeling; @ascarza_retention_2018},
in a contractual business the customer needs to renew their contracts to continue its usage
[@Ascarza_Hardie_2013], against non contractual where the firm has to infer if 
the customer is still active
However, in contractual settings the customer dropout represents an explicit
ending of a relationship which is more penalizing than non contractual settings
[@risselada_staying_2010]. 
This has implications to the profitability of the organizations increasing marketing 
costs and reducing sales [@amin_customer_2017].

The anticipation of the dropout allows the development of countermeasures to
reduce customer churn. Several studies address the problem related to customer
retention trying to improve the profitability [@coussement_improving_2009;
@garcia_intelligent_2017; @devriendt_why_2019].
Existing organizations are addressing this problem by shifting their target from 
capturing new customers to preserving existing ones [@garcia_intelligent_2017],
considering that investments in retention strategies are more profitable 
than acquiring new customers [@coussement_improving_2009].

The approaches normally employed use a dependent variable representing dropout
or non-dropout, without considering a dynamic perspective that the dropout risk
changes over time [@Alboukaey_dynamic_2020]. The survival models try to solve
this limitation [@routh_estimating_2020] capturing a temporal dimension of the
customer dropout [@perianez_churn_2016]. @perianez_churn_2016 used survival
analysis to predict also when the dropout will occur.

Other studies proposed also the integration of several algorithms to improve the
performance in the prediction of the dropout such the usage of clusters combined
with churn prediction [@hung_applying_2006; @gok_case_2015;
@vijaya_sivasankar_2019]. The approach relies in the assumption that combining
the customers in different clusters allows the improvement of the prediction
accuracy. @vijaya_sivasankar_2019 suggested the adoption hybrid models combining
more than one classier are achieving increased performance compared to those
using single classifiers.

There are several challenges around the timing related to dropout, or
considering the dynamic behavior of the customer in the intent to drop out
[@alboukaey_dynamic_2020]. The importance of understanding when dropout will
occur and the risk when discarding the temporal perspective of the problem seems
to be an element that should be addressed. Few studies considered this
[@perianez_churn_2016; @burez_separating_2008]. This shows an opportunity to
address the importance of the timeframe and its influence on the efficiency of
the model and also evalute if the combination of clusters could improve the
performance.

Survival analysis, which origin as stands in biomedical statistics, its especially 
well-suited to studying the timing of events in longitudinal data [@Singer_Willett_1993]. 
Survival analysis is a class of statistical methods modelling the occurrence and timing of 
an event, such as the customer dropout. 
Survival analysis allow us to examine not only if an event occurred but also how long it 
took to occur. 
A primary value of survival analysis, however, is to compare dropout probability for 
individuals classified with theoretically relevant variables. 
The survival methods have enjoyed and increasing popularity in several disciplines ranging 
from medicine to economics [@Singer_Willett_1993].

Random Survival Forests does not make the proportional hazards assumption
[@Ehrlinger_2016] and has the flexibility to model survivor curves that are of
dissimilar shapes for contrasting groups of subjects. Random Survival Forest is
an extension of Random Forest allowing efficient non-parametric analysis of time
to event data [@Breiman_2001]. This characteristics allow us to surpass the Cox
Regression limitation of the proportional hazard assumption, requiring to
exclude variables which not fulfill the model assumption. It was shown by
@Breiman_2001 that ensemble learning can be further improved by injecting
randomization into the base learning process - a method called Random Forests.

Some studies employed also a combination of clusters with the churn prediction 
[@hung_applying_2006;@gok_case_2015;@vijaya_sivasankar_2019], where the customers
where grouped in clusters to improve the accuracy within each cluster.
Clusters are approaches using unsupervised algorithms to group elements
with similar characteristics. 
This unsupervised method has been widely used employing approaches
such Hierarchical Clustering [@Saunders_1980], K-Means [@vijaya_sivasankar_2019], or
Random Forest Clustering [@Breiman_2001].
@JafariMarandi_Denton_Idris_Smith_Keramati_2020 explored an approach combining
clustering methods in parallel to a classification.

In this study, we investigate if an hybrid approach using clusters and random survival forests 
which have never been used in to predict membership in a health club using existing data 
improves the prediction accuracy. 
Our paper is organized as follows. In the next section, we address our research methodology,
using survival analysis to identify the survival probability along the time, how 
we address the problem of determining the clusters and the performance of the model 
to predict customer survival within each identified cluster.

# Survival analysis

Survival analysis focus in the analysis of the time until an event of interest,
and exploring its relationship with different factors.
The main advantage is related to the concept of censoring, indicating
that observations that are not complete related to the event of interest, e.g.
customers that didn't dropout yet, which are incorporated in the analysis.
This means that there are customers still active for which we don't know if 
the event of dropout has occurred, which is called censorship.
Survival models take censoring into account and incorporate this uncertainty, 
instead of predicting the time of event such in regression models, the survival models
allow to predict the probability of an event happens at a particular time.

The time of dropout is represented by T, which is a non-negative random variable, 
indicating the time period of the event occurring for a randomly selected individual 
from the population, representing the probability of an event to occur each time period given 
that has not already occurred in a previous time period, known as discrete-time hazard function 
[@Singer_Willett_1993]. 
The survival function represents the probability of an individual surviving after time t, 
S(t) = P(T $>$ t), t $\geq$ 0, with the properties S(0) = 1,S($\infty$) = 0. 
The distribution function is represented with F, defined as F(t) = P(T $\leq$ 0), for t $\geq$ 0. 
The function of probability density represented with f where:

\begin{equation}
f(t) = \lim_{dt{\to 0}} \frac{P[t \le T < t + dt ]}{dt}
\end{equation}

$f(t)dt$ represents the probability of an event occurring in the moment t. 
The need to represent the distribution evolution of the death probability along the time, 
uses to the hazard function, represented as:

\begin{equation}
\lambda(t) = \lim_{dt \to 0} \frac{P[t \le T < t + dt | T \ge t ]}{dt}
\end{equation}

The determination of the survival curves is based in the following elements: 
(1) the total value of observations removed during the time period (e.g. days, months or years), 
    either by dropout or by censorship; 
(2) observations that composed the sample of the study; 
(3) customers who had not yet dropped out at any given time. 
The survival probability until the time period ii ($p_i$) is calculated with:

\begin{equation}
p_i=\frac{r_i-d_i}{r_i}
\end{equation}

Where $r_i$ is the number of individuals that survived at the beginning of the period, 
$d_i$ the number of individuals who left during the period. 
The survival time estimate was also taken considering the month in which it is found (estimated). 
Cox's allow test difference between survival times. 
The advantage in using survival analysis was that allow us to detect if the risk of an event differs 
systematically across different people, using specific predictors. 
The coefficients in a Cox regression were related to the hazard, where a positive value 
represents a worse prognosis and the opposite, negative value a better prognosis.
The advantage of survival analysis was that allow us to include information of covariates
that were censored up to the censoring event.

The Cox PH model assumes the covariates to be time independent, for example
gender and age when where retrieved do not change over time [@Schober2018Survival]
Because the Cox model requires the hazards in both groups to be proportional, 
researchers are often asked to "test" whether hazards are proportional
[@Stensrud_Hernán_2020].
Considering this we explored other approach that allow us to develop this analysis without 
the proportional hazard assumptions, the survival trees.

# Survival Trees

Survival trees are methods based in tree based models based on Random Forest [@Breiman_2001].
Random survival forests is an ensemble method for analysis of right-censored
data [@Ishwaran_Kogalur_Blackstone_Lauer_2008], using randomization to 
improve the performance.
Random survival forests follows this framework [@Ishwaran_Kogalur_Blackstone_Lauer_2008]:

1. Draw B random samples of the same size from the original dataset with replacement. The samples that are not drawn are said to be out-of-bag (OOB). Grow a survival tree on each of the b = 1 , . . . , B samples.
2. At each node, select a random subset of predictor variables and find the best predictor and splitting value that provide two subsets (the daughter nodes) which maximizes the difference in the objective function.
3. Repeat step 2 recursively on each daughter node until a stopping criterion is met.
4. Calculate a cumulative hazard function (CHF) for each tree and average over all CHFs for the B trees to obtain the ensemble CHF.
5. Compute the prediction error for the ensemble CHF using only the OOB data.

In each node is selected a predictor $x$ from a random selected predicted variables
and split value $c$ (one unique value of $x$).
Each sample $i$ if assigned the daughter right node if $x_i \le c$ or left if 
$x_i \ge c$, then is calculated the logrank such as:

\begin{equation}
L(x, c) = \frac{ \sum^{N}_{i=1} \left(  d_{i, 1} - Y_{i,1} \frac{d_i}{Y_i} \right)  }
               { \sqrt{  \sum^{N}_{i=1}  \frac{Y_{i,1}}{Y_i} \left( 1 - \frac{Y_{i,1}}{Y_i} \right) \left( \frac{Y_i-d_i}{Y_i-1} \right) d_i  } }
\end{equation}

Where:

* $j$: Daughter node, $j \in \{1,2\}$
* $d_{i,j}$: Number of events at time $t_i$ in daughter node $j$
* $Y_{i,j}$: Number of elements that had the event or are in risk at time $t_i$
           in daughter node $j$
* $d_i$: Number of events at time $t_i$, such $d_i=\sum_j{d_{i,j}}$
* $Y_i$: Number of elements that experienced an event or are at risk at
      $t_i$ so $Y_i=\sum_j Y_{i,j}$

We loop every $x$ and $c$ until find $x^*$ that satisfy $|L(x^{*}, c^{*})| \geq |L(x, c)|$
for every $x$ and $c$.
The model performance was determined with the concordance probability (C-index),
Brier Score (BS) and Mean Absolute Error (MAE) [@wangmachine2017]. The feature
importance was determined calculating the difference between the true class
label and noised data [@Breiman_2001].

The BS is used to evaluate the predicted accuracy of the survival function
at a given time $t$. 
Representing the average square distance between the survival status and the
predicted survival probability, where the value 0 is the best possible outcome.

\begin{equation}
BS(t) =  \frac{1}{N} \sum_{i = 1}^{N} \left( \frac{\left( 0 - \hat{S}(t, \vec{x}_i)\right)^2 \cdot \mathbb{1}_{T_i \leq t, \delta_i = 1}}{ \hat{G}(T_i^-)} + \frac{ \left( 1 - \hat{S}(t, \vec{x}_i)\right)^2 \cdot \mathbb{1}_{T_i > t}}{ \hat{G}(t)} \right)
\end{equation}

The model should have a Brier score below $0.25$. 
Considering that if $\forall i \in [\![1, N]\!], \hat{S}(t, \vec{x}_i) = 0.5$, then $BS(t) = 0.25$.

# Methodology

To simplify the analysis, the survival probabilities are presented as a survival curve. 
The survival curve is a representation of the survival probabilities corresponding to a 
time where the events are observed [@Bland_Altman_1998].
The survival analysis was conducted using the package Lifelines [@Pilon_2021].
Where dropout is a binary value where one represent churn and zero not churn. The
dropout happens when a member does not have a payment.

The random survival forest was developed using the package PySurvival
[@Fotso_others_2019]. PySurvival is an open source python package for 
Survival Analysis modeling - the modeling concept used to analyze or predict 
when an event is likely to happen.
The model was built with with 70\% of the data for training and 30\% for testing.

The model performance was determined with the concordance probability (C-index),
Brier Score (BS) and Mean Absolute Error (MAE) [@wangmachine2017]. The feature
importance was determined calculating the difference between the true class
label and noised data [@Breiman_2001].

The BS measure the average discrepancies between the status (dropout/non-dropout)
and the estimated probabilities at a given time.
The Integrated Brier Score (IBS) was used to calculate the performance 
in all available times (from $t_1$ to $t_{max}$) as:

\begin{equation}
IBS = \int_{t_1}^{t_\text{max}} {BS}^c(t) d w(t)
\end{equation}

Representing the average square distance between the survival status and the
predicted survival probability, where the value 0 is the best possible outcome.

The calculation of he number of clusters used the package mclust [@scrucca2016]
using the Bayesian Information Criterion (BIC). The model that gives the minimum
BIC score can be selected as the best model [@schwarz1978] simplifying the
problem related to choosing the number of components and identifying the
structure of the covariance matrix, based on modelling with multivariate normal
distributions for each component that forms the data set [@akogul2016].

The hybrid approach was develop as follows:

1. Identify the optimal number of clusters using [@scrucca2016]
2. Fit the model using the identified number of clusters
3. Estimate for each element the cluster 
4. for each cluster follow the framework proposed by @Ishwaran_Kogalur_Blackstone_Lauer_2008
   to calculate the random survival model

# Dataset

In this study, data from 5,209 fitness customers was analysed (mean age = 27.88, SD=11.80 years) 
from a Portuguese fitness centre. 
The data was collected from software e\@sport (Cedis, Portugal) between 2014 and 2017. 
The information retrieved was: Age of the participants in years; Sex (0-female, 1-male); 
Non-attendance days before dropout; Total amount billed; Average number of visits per week;
Total number of visits; Weekly contracted accesses; Number of registration renewals; 
Number of customer referrals; Registration month; Customer enrolment duration; and status 
(dropout/non-dropout).
Dropout event occur when customer communicate the intention to terminate the contract or 
did not pay the monthly fee during 60 days.

```{r get_data, echo = FALSE, message = FALSE, warning=FALSE, include = FALSE}
library(stargazer)
library(readxl)
library(dplyr)
library(visdat)

df_members <- read_excel("data/fitness_customers.xlsx")

# rename column labels
names(df_members) <- c("id", "age", "sex", "dayswfreq", "tbilled",
                       "maccess", "freeuse", "nentries", "cfreq", "nrenewals",
                       "cref", "start_date", "months", "dropout")

# remove null values
df_members <- df_members[complete.cases(df_members), ]

names(df_members)

# select relevant variables
df_members <- df_members %>%
select(age, sex, dayswfreq, tbilled, maccess, freeuse,
       nentries, cfreq, months, dropout)

str(df_members)
vis_dat(df_members) #check
```

Table \@ref(tab:summarytable) shows data's summary statistics. The average age
is `r round(mean(df_members$age,na.rm = TRUE),1)` ±
`r round(sd(df_members$age,na.rm = TRUE),1)`, the entries are 
`r round(mean(df_members$nentries,na.rm = TRUE),0)` ±
`r  round(sd(df_members$nentries,na.rm = TRUE),1)` with an inscription period of
`r round(mean(df_members$months,na.rm = TRUE),0)` ±
`r  round(sd(df_members$months,na.rm = TRUE),1)` months.

```{r teste, eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE, paged.print=FALSE}
library(stargazer)
df_summary <- summary(df_members)
stargazer(df_members,
          title = "Summary statistics",
          label = "tab1cars",
          table.placement = "h",
          header = FALSE,
          summary = TRUE,
          summary.stat = c("min", "p25", "median", "p75", "max", "median", "sd")
         )
```

```{r summarytable, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(gtsummary)
library(kableExtra)
library(labelled)

var_label(df_members$age) <- "Age in years"
var_label(df_members$sex) <- "Male or female"

tbl <- df_members %>%
    tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})",
                                 all_categorical() ~ "{p}%"),
                type = list(age ~ "continuous")) %>%
    add_stat_label()

as_kable_extra(tbl, booktabs = T,
               caption = "Summary statistics of features used")
```
Figure \@ref(fig:membershipmonth) shows the distribution of the dropout
considering the number of months of membership.

```{r, label="membershipmonth", eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Number of members by month"}
members_month <- df_members %>%
  select(months, dropout) %>%
  group_by(months, dropout) %>%
  summarize(count = n())

members_month$dropout <- factor(members_month$dropout)
g <- ggplot(data = members_month,
            mapping = aes(x = months, y = count, linetype = dropout))
g <- g + geom_line() + labs(x = "months", y = "Number of members") +
    theme_classic()
g

# Export the file 
# ggsave(file="/mnt/c/nuvem/Dropbox/doutoramento/tese/9.tese_documento/PhD-Pedro-Sobreiro/figures/membershipyear.pdf",device="pdf")
```

```{python correlationMatrix, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Correlation matrix"}
from pysurvival.utils.display import correlation_matrix
import matplotlib.pyplot as plt
import pandas as pd
import numpy as np

col = ['sex']

df_members = r.df_members #copy r dataframe to python

# convert to int
df_members['age']=df_members['age'].astype(int)
df_members['dayswfreq']=df_members['dayswfreq'].astype(int)
df_members['cfreq']=df_members['cfreq'].astype(int)
df_members['months']=df_members['months'].astype(int)
df_members['dropout']=df_members['dropout'].astype(int)
df_members['sex']=df_members['sex'].astype(int)

df_members = pd.get_dummies(df_members, columns=col,drop_first=True)

# Creating the time and event columns
time_column = 'months'
event_column = 'dropout'

# Extracting the features
features = np.setdiff1d(df_members.columns, [time_column, event_column] ).tolist()

correlation_matrix(df_members[features], figure_size=(10,10), text_fontsize=6)

r.df_members = df_members

#plt.plot()
```

# Results

The table \@ref(tab:survivalprobabilities) depicts the data of the survival time of the customers 
during the first months, the results showed that the customers have a survival probability of 24.44\% 
at 12 months (column $p_i$ - likelihood probability) with a median survival time of 10 months
(column estimated_survival).
The survival probability at 6 months was 54.5\%, representing an risk of dropout of 
45.5\% with a estimated survival of 6 months.

```{python lifelines, include=FALSE}
from lifelines import KaplanMeierFitter
kmf = KaplanMeierFitter()
T = df_members['months']
C = df_members['dropout']

kmf.fit(T,C,label="Customers")

kmf.event_table.reset_index()
kmf.conditional_time_to_event_

survival_table = pd.concat([kmf.event_table.reset_index(),
                            kmf.conditional_time_to_event_.reset_index(),
                            kmf.survival_function_.reset_index()],axis=1)

survival_table.drop(['timeline'],axis=1,inplace=True)
survival_table.columns = ['event_at', 'removed', 'observed', 'censored', 
                          'entrance', 'at_risk', 'estimated_survival', 
                          'prob']
```

```{r survivalprobabilities, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(gtsummary)
library(kableExtra)
library(labelled)
survival_table <- py$survival_table
var_label(survival_table$event_at) <- "Event Month"
var_label(survival_table$removed) <- "Removed"
var_label(survival_table$observed) <- "Dropout"
var_label(survival_table$censored) <- "Censored"
var_label(survival_table$at_risk) <- "Risk of dropout"
var_label(survival_table$estimated_survival) <- "Estimated survival (months)"
var_label(survival_table$prob) <- "Survival Probability"

tbl <- head(survival_table, 25)

kbl(tbl, booktabs = T,
    format = "latex",
    caption = "Determination of the survival time probabilities",
    digits = 3) %>%
    footnote(general = paste("Removed – the sum of customers with dropout and that are",
                             "censored; Censored – the event did not occur during the",
                             "period of this data, collection; Risk of Dropout – ",
                             "number of customers at risk of, dropout; pi – survival",
                             "probability; Estimated Survival - months to survive in",
                             "the sports facility.", sep = " "),
                              threeparttable = TRUE)
```

Figure \@ref(fig:lifelinesplot) shows the Kaplan Meier survival curve customers 
considering the number of months of membership (x axis) and survival probability 
(y axis).
The customer dropout is very high in the first 12 months, ranging from a survival
probability of 54\% after the first 6 months until 24\% after 12 months.

```{python lifelinesplot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Survival probabilities"}
import matplotlib.pyplot as plt
import matplotlib.font_manager as font_manager

plt.rcParams['figure.figsize'] = [12, 7]

fontL = font_manager.FontProperties(family='Times New Roman',weight=None,style='normal', size=10)

# fontAxis = {'family': 'Times New Roman'}

ax = kmf.plot(color='black')
ax.legend(prop=fontL)
#ax.set_xlabel('Months',fontdict=fontAxis,fontsize=12)
ax.set_xlabel('Months')
#ax.set_ylabel('Survival probability',fontdict=fontAxis,fontsize=12)
ax.set_ylabel('Survival probability')

ax.axvline(x=6,ymax=0.54,linestyle='--',color='black');ax.axhline(y=0.54,xmax=0.14,linestyle='--',color='black')
ax.annotate("0.54",xy=(6, 0.54), xytext=(7, 0.7))

ax.axvline(x=12,ymax=0.24,linestyle='--',color='black');ax.axhline(y=0.24,xmax=0.254,linestyle='--',color='black')
ax.annotate("0.24",xy=(12, 0.24), xytext=(12, 0.40))

ax.axvline(x=18,ymax=0.15,linestyle='--',color='black');ax.axhline(y=0.15,xmax=0.38,linestyle='--',color='black')
ax.annotate("0.15",xy=(18, 0.15), xytext=(18, 0.30))

plt.savefig(r.caminho_figuras+'fitnessLifelinesPlot.png', dpi=300)
plt.show()

plt.close()
```

Figure \@ref(fig:logrank) shows the survival by gender.
The survival curves by gender are very similar, both types of 
customers present a behavior that is not very different.

```{python logrank, results='hide', echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Survival by gender"}
import matplotlib.pyplot as plt

ax = plt.subplot(111)

plt.rcParams['figure.figsize'] = [12, 7]

for gen in df_members['sex_1'].unique():
    ix = df_members['sex_1'] == gen
    kmf.fit(T.loc[ix], C.loc[ix], label=str(gen))
    ax = kmf.plot(ax=ax)

plt.savefig(r.caminho_figuras+'fitnessLogrankGender.png', dpi=300)

plt.show()
plt.close()
```

Figure \@ref(fig:logrankcfreq) shows the survival by contracted frequency.
Customers with contracted frequency of 6 and 4 times a week have higher survival
probabilities, against lower of customers with contracted frequencies of 7 and 
2 times a week.
Survival curves allow to explore tendencies related to survival to extract
actionable knowledge.

```{python logrankcfreq, results='hide', echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Survival by contracted frequency"}
import matplotlib.pyplot as plt

ax = plt.subplot(111)

plt.rcParams['figure.figsize'] = [12, 7]

for item in df_members['cfreq'].unique():
    ix = df_members['cfreq'] == item
    kmf.fit(T.loc[ix], C.loc[ix], label=str(item))
    ax = kmf.plot(ax=ax)

plt.savefig(r.caminho_figuras+'fitnessLogrankCfreq.png', dpi=300)

plt.show()
plt.close()
```

```{python coxRegression, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Cox regression"}
vars= ['age', 'dayswfreq', 'tbilled', 'maccess', 'freeuse', 'nentries',
       'cfreq', 'months', 'sex_1', 'dropout']
df_regression = df_members[vars].copy()

from lifelines import CoxPHFitter
cph = CoxPHFitter()
cph.fit(df_regression,duration_col='months',event_col='dropout')

cph.check_assumptions(df_regression)
```

The proportional hazard assumptions failed in the following variables:
age p<0.01, cfreq p<0.01, dayswfreq p<0.01, tbilled p<0.01, freeuse 
p<0.01, nentries p<0.01.

## Survival trees

To evaluate the performance of the random survival forest we have calculated
the concordance probability (C-index), IBS and Mean Absolute Error (MAE).
The IBS presents an accuracy along the 12 months of 0.08 (figure \@ref(fig:createModel1)).

```{python createModel1, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Model performance"}
from pysurvival.models.survival_forest import RandomSurvivalForestModel
from sklearn.model_selection import train_test_split
from pysurvival.utils.metrics import concordance_index
from pysurvival.utils.display import integrated_brier_score
from pysurvival.utils.display import compare_to_actual

X = df_members.copy()
t = df_members['months']
e = df_members['dropout']
X.drop(axis=1, columns=['months','dropout'],inplace=True)

X_train, X_test, t_train, t_test, e_train, e_test = train_test_split(X, t, e, test_size=0.3, random_state=0)

# Fitting the model
csf = RandomSurvivalForestModel(num_trees=20)
csf.fit(X_train, t_train, e_train, max_features='sqrt', max_depth=5, min_node_size=20, seed = 1)

c_index = concordance_index(csf, X_test, t_test, e_test)
ibs = integrated_brier_score(csf, X_test, t_test, e_test, t_max=12, figure_size=(12,5))
#plt.savefig(r.caminho_figuras+'create_model1_fitness.png', dpi=300)
```

The actual versus predicted model presents the actual and predicted customers which dropout 
during the 40 months, which as an average absolute error of 7.5 customers (figure \@ref(fig:createModel2)).

```{python createModel2, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Conditional survival forest"}

results = compare_to_actual(csf, X_test, t_test, e_test, is_at_risk = False,  figure_size=(12, 6), metrics = ['rmse', 'mean', 'median'])
```

```{r previsaoglobal, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
# gravar a figura com a previsão global
#ggsave(file="/mnt/c/nuvem/Dropbox/doutoramento/tese/9.tese_documento/PhD-Pedro-Sobreiro/figures/previsao_global.pdf",device="pdf")

```

Table \@ref(tab:summarytable2) shows features importance calculated according [@Breiman_2001],
where the percent increase in misclassification rate as compared to the out-of-bag rate (with all 
variables intact), out-of-bag is a bootstrap aggregating (subsampling with replacement to create 
training samples for the model to learn from) where two independent sets are created. 
One set, the bootstrap sample, data chosen to be "in-the-bag" by sampling with replacement and the 
out-of-bag is all data not chosen in the sampling process.
The most important variable is the $dayswfreq$, followed by $tbilled$ and
$nentries$, compared with the $cfreq$, $age$, and $sex$.

```{r summarytable2, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
tbl <- py$csf$variable_importance_table
kbl(tbl, booktabs = T, caption = "Features importance in the survival model")
```

The prediction is very similar to the actual value. 
The model accuracy is very high with a root mean square error of
`r round(py$results$root_mean_squared_error)`.
The mean absolute error mean was `r round(py$results$mean_absolute_error,2)` customers,
and the median absolute error was `r round(py$results$median_absolute_error,2)`.

## Survival trees based model with clusters

In this approach we have created clusters and applied the survival trees within 
each cluster.
The determination of the clusters using the BIC criterion where the EEV model: 7 clusters -57159.24; 
6 clusters -63937.59; and 4 clusters -77088.81 figure \@ref(fig:clusters) shows the 
determination of the number of clusters using BIC, also the elbow analysis available in 
figure \@ref(fig:elbowCalculation). 
An optimal number of clusters was considered of five. 
Considering that was the value after the average distortion was flattened.

```{r clusters, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Analysis number of clusters"}
library(mclust)
#y <- scale(py$df_members)
y <- scale(py$df_members)

set.seed(0) # to make reproducible

#bic <- mclustBIC(y,modelNames = c("VEI"))
bic <- mclustBIC(y)
# Best model using the BIC criteria
#bic
plot(bic, what = "BIC")

```

```{r NbClust, echo = FALSE, message = FALSE, warning = FALSE, eval=FALSE}
library(NbClust)
nb <- NbClust(y, diss = NULL, distance = "euclidean",
              min.nc = 2, max.nc = 5, method = "kmeans",
              index = "all", alphaBeale = 0.1)
hist(nb$Best.nc[1, ], breaks = max(na.omit(nb$Best.nc[1, ])))

```

```{python elbowCalculation, results='hide', echo=FALSE, message=FALSE, warning = FALSE, fig.cap="Elbow analysis"}
from sklearn.cluster import KMeans
from sklearn import preprocessing
from scipy.spatial.distance import cdist
import matplotlib.pyplot as plt
import numpy as np

#Finding optimal no. of clusters
clusters=range(1,20)
meanDistortions=[]

for k in clusters:
    model=KMeans(n_clusters=k)
    model.fit(df_members)
    prediction=model.predict(df_members)
    meanDistortions.append(sum(np.min(cdist(df_members, model.cluster_centers_, 'euclidean'), axis=1)) / df_members.shape[0])

plt.plot(clusters, meanDistortions, 'bx-')
plt.xlabel('k')
plt.ylabel('Average distortion')
plt.show();

```

The calculation of the clusters to each member in the dataset was developed considering,
7,6 and 4 clusters. 
However after the executions KMeans only performed with 3 clusters.

```{python kmeanscalculation, results='hide', echo=FALSE, message=FALSE, warning=FALSE}
# we are going to use random_state = 0 for the centroid
# initialization being deterministic allowing a better
# reproducibility
cluster = KMeans(n_clusters=3, random_state=0)

cluster.fit(df_members)
df_members['cluster']=cluster.predict(df_members)
#print(df_members.cluster.value_counts());
```

```{python kmeans PCA, echo=FALSE, message = FALSE, warning = FALSE}
from sklearn.decomposition import PCA

plt.rcParams['figure.figsize'] = [13, 4]

pca=PCA(n_components=2)
df_members['X_pca']=pca.fit_transform(df_members)[:,0] #todas as linhas da primeira coluna com redução
df_members['Y_pca']=pca.fit_transform(df_members)[:,1] #todas as linhas da segunda coluna com redução

fig, ax = plt.subplots()
for cluster in df_members.cluster.unique():
    x = df_members['X_pca'].loc[df_members.cluster == cluster]
    y = df_members['Y_pca'].loc[df_members.cluster == cluster]
    ax.scatter(x, y, label=cluster, alpha=1, edgecolors='none');

ax.legend()
ax.grid(True)
ax.set_title('Customers clusters');
plt.show();
```

The model accuracy is very high in the first years. The prediction is very
similar to the actual value. The absolute error mean of
6 customers.

```{python createModelClusters, results='hide', include=TRUE, echo=FALSE, message = FALSE, warning=FALSE}
# Building training and testing sets for the clusters

df_resultados = pd.DataFrame(columns=['cluster','rmse','mean','median'])

for cluster in df_members.cluster.unique():
    # Number of samples in the dataset
    df_members_cluster = df_members[df_members.cluster == cluster].copy()
    X = df_members_cluster.copy()
    t = df_members_cluster['months']
    e = df_members_cluster['dropout']
    X.drop(axis=1, columns=['months','dropout'], inplace=True)
    X_train, X_test, t_train, t_test, e_train, e_test = train_test_split(X, t, 
                                                        e, random_state=0)
    # Fitting the model
    csf = RandomSurvivalForestModel(num_trees=20)
    csf.fit(X_train, t_train, e_train, max_features='sqrt', max_depth=5, 
            min_node_size=20, seed = 1)

    results_cluster = compare_to_actual(csf, X_test, t_test, e_test, 
                                is_at_risk = False, figure_size=(12, 6), 
                                metrics = ['rmse', 'mean', 'median'])

    #print(f"The cluster {cluster} as a size of {X.shape[0]}")
    #print(results_cluster)
    #print(f"Cluster {cluster} RMSE {results_cluster['root_mean_squared_error']}")
    #print(f"Cluster {cluster} MAE {results_cluster['median_absolute_error']}")
    #print(f"Cluster {cluster} MAError {results_cluster['mean_absolute_error']}")

    df_resultados = df_resultados.append({'cluster' : cluster, 
                                          'rmse' : results_cluster['root_mean_squared_error'], 
                                          'mean' : results_cluster['median_absolute_error'], 
                                          'median' : results_cluster['mean_absolute_error']}, 
                                          ignore_index = True)
```

The performance of the cluster 1 the IBS presents an accuracy of 0.06 
(figure \@ref(fig:createModelCluster1IBS)) along all time.
The actual versus predicted model presents the actual and predicted customers which dropout during
the 40 months, which as an mean absolute error of 1.5 customers, the mean median absolute
error was 0.61 and the Root Mean Square Error of 2.8 (figure \@ref(fig:createModelCluster1SF)).

The features importance in the survival model cluster 1 (table \@ref(tab:summarytableCluster1))
identify the three most relevant features to predict survival $maccess$, 
$tbilled$, and $dayswfreq$. The features with lower relevance were $freeuse$, $sex$ and $cfreq$.

```{python createModelCluster1IBS, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Model performance cluster 1"}
df_members_cluster = df_members[df_members.cluster == 0].copy()
X = df_members_cluster.copy()
t = df_members_cluster['months']
e = df_members_cluster['dropout']
X.drop(axis=1, columns=['months','dropout','X_pca','Y_pca','cluster'], inplace=True)
X_train, X_test, t_train, t_test, e_train, e_test = train_test_split(X, t, 
                                                    e, random_state=0)
# Fitting the model
csf = RandomSurvivalForestModel(num_trees=20)
csf.fit(X_train, t_train, e_train, max_features='sqrt', max_depth=5, 
        min_node_size=20, seed = 1)

c_index = concordance_index(csf, X_test, t_test, e_test)
ibs = integrated_brier_score(csf, X_test, t_test, e_test, t_max=12, 
                             figure_size=(12,5))

```

```{python createModelCluster1SF, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Conditional survival forest cluster 1"}
compare_to_actual(csf, X_test, t_test, e_test, is_at_risk = False, figure_size=(12, 6), 
                  metrics = ['rmse', 'mean', 'median'])

```

```{r summarytableCluster1, echo=FALSE, message=FALSE, warning=FALSE}
tbl <- py$csf$variable_importance_table
kbl(tbl, booktabs = T,
    caption = "Features importance in the survival model with cluster 1")
```

The performance of the cluster 2 the IBS presents an accuracy along time 0.09 
(figure \@ref(fig:createModelCluster2IBS)) along all time.
The actual versus predicted model presents the actual and predicted customers which dropout during
the 40 months, which as an mean absolute error of 6.5 customers, the mean median absolute
error was 2.3 and the Root Mean Square Error of 11.79 (figure \@ref(fig:createModelCluster2SF)).
The features importance in the survival model cluster 2 (table \@ref(tab:summarytableCluster2))
identify the three most relevant features to predict survival $dayswfreq$, 
$tbilled$, and $freeuse$. The least relevant were $nentries$, $cfreq$, and $age$.

```{python createModelCluster2IBS, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Model performance cluster 2"}
df_members_cluster = df_members[df_members.cluster == 1].copy()
X = df_members_cluster.copy()
t = df_members_cluster['months']
e = df_members_cluster['dropout']
X.drop(axis=1, columns=['months','dropout','X_pca','Y_pca','cluster'], inplace=True)
X_train, X_test, t_train, t_test, e_train, e_test = train_test_split(X, t, 
                                                    e, random_state=0)
# Fitting the model
csf = RandomSurvivalForestModel(num_trees=20)
csf.fit(X_train, t_train, e_train, max_features='sqrt', max_depth=5, 
        min_node_size=20, seed = 1)

c_index = concordance_index(csf, X_test, t_test, e_test)
ibs = integrated_brier_score(csf, X_test, t_test, e_test, t_max=12, 
                             figure_size=(12,5))

```

```{python createModelCluster2SF, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Conditional survival forest cluster 2"}
compare_to_actual(csf, X_test, t_test, e_test, is_at_risk = False, figure_size=(12, 6), 
                  metrics = ['rmse', 'mean', 'median'])

```

The performance of the cluster 3 the IBS presents an accuracy along time 0.01 
(figure \@ref(fig:createModelCluster3IBS)) along all time.
The actual versus predicted model presents the actual and predicted customers which dropout during
the 40 months, which as an mean absolute error of 1.5 customers, the mean median absolute
error was 1.2 and the Root Mean Square Error of 2.08 (figure \@ref(fig:createModelCluster3SF)).
The features importance in the survival model cluster 3 (table \@ref(tab:summarytableCluster3))
identify the three most relevant features to predict survival $tbilled$, 
$dayswfreq$, and $nentries$. 
The least relevant were $sex$, $age$, and $cfreq$.

```{r summarytableCluster2, echo=FALSE, message=FALSE, warning=FALSE}
tbl <- py$csf$variable_importance_table
kbl(tbl, booktabs = T,
    caption = "Features importance in the survival model with cluster 2")
```

```{python createModelCluster3IBS, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Model performance cluster 3"}
df_members_cluster = df_members[df_members.cluster == 2].copy()
X = df_members_cluster.copy()
t = df_members_cluster['months']
e = df_members_cluster['dropout']
X.drop(axis=1, columns=['months','dropout','X_pca','Y_pca','cluster'], inplace=True)
X_train, X_test, t_train, t_test, e_train, e_test = train_test_split(X, t, 
                                                    e, random_state=0)
# Fitting the model
csf = RandomSurvivalForestModel(num_trees=20)
csf.fit(X_train, t_train, e_train, max_features='sqrt', max_depth=5, 
        min_node_size=20, seed = 1)

c_index = concordance_index(csf, X_test, t_test, e_test)
ibs = integrated_brier_score(csf, X_test, t_test, e_test, t_max=12, 
                             figure_size=(12,5))

```

```{python createModelCluster3SF, results='hide', echo=FALSE, message = FALSE, warning = FALSE, fig.cap = "Conditional survival forest cluster 3"}
compare_to_actual(csf, X_test, t_test, e_test, is_at_risk = False, figure_size=(12, 6), 
                  metrics = ['rmse', 'mean', 'median'])

```

The features importance in the survival model (table \@ref(tab:summarytableCluster3))
identify the three most relevant features to predict survival $dayswfreq$, 
$tbilled$, and $nentries$.

```{r summarytableCluster3, echo=FALSE, message=FALSE, warning=FALSE}
tbl <- py$csf$variable_importance_table
kbl(tbl, booktabs = T,
    caption = "Features importance in the survival model with cluster 3")
```

### Model Comparison

Table \@ref(tab:summarytable3) shows the performance of both approaches, with or without clusters. 
The RMSE in the clusters 1, 2 and 3 is lower than not using clusters to predict dropout.
Overall the performance improved. The performance is also better using mean and median.

```{r summarytable3, echo=FALSE, message=FALSE, warning=FALSE}
tbl <- py$df_resultados
# append results without clusters
tbl <- rbind(tbl, c("w/cluster",
                   py$results$root_mean_squared,
                   py$results$median_absolute_error,
                   py$results$mean_absolute_error))
kbl(tbl, booktabs = T,
    caption = "Performance of prediction in each cluster", digits = 3)
```

The model accuracy without clusters is very high with a root mean
square error of 13, the mean absolute error mean was 7.53 customers, and the median 
absolute error was 4.04.
The model using clusters had an mean absolute error of 1.5 customers, the mean median absolute
error was 1.2 and the Root Mean Square Error of 2.08.
The performance using clusters improved significantly.

# Conclusion

This paper investigated the customer dropout in a Health Club organization, using
a dynamic perspective that the dropout risk varies along the time.
Exploring two approaches, using a survival model based on random forests 
with or without clusters.
The model using clusters allowed to combine the customers in different clusters,
an hybrid approach.
Based on this performance the proposed model using clusters allows to improve 
the accuracy on the survival model allowing to target approaches considering
the timing when the dropout occurs, considering the clusters where the customer is.
Is very important for managers use this information to improve their retention
strategies.

<!--

```{=tex}
\begin{itemize}
    \item RQ1: What is the current state of the research being developed?
    \item RQ2: What algorithms have been used to predict dropout?
    \item RQ3: What are the features used to predict dropout?
    \item RQ4: When does dropout occur?
    \item RQ5: How is the accuracy of the machine learning algorithms in predicting dropout measured?
\end{itemize}
```
From RQ1, it was possible to identify some business areas that are
under-researched, such as the energy sector, education, logistics and
hospitality. Compared to other business areas such telecom or the financial
sector, research on the energy sector or water supply is lacking, considering
the contractual settings that are assumed to provide such types of services.
Considering the business model of many software companies as software as a
services (SaS), the number of research works is also surprisingly low.

RQ2 also provided an overall perspective related to the algorithms being used to
predict customer dropout. The first could be the importance and wider adoption
of decision trees and random forests [@antipov_applying_2010;
@benoit_improving_2012; @burez_crm_2007], and logistic regression
[@coussement_improved_2010], which could be due to its higher interpretability
and flexibility [@keramati_improved_2014]. Interpretability is an important
aspect for the marketing department in the extraction of valuable information
from the model to develop effective retention strategies [@verbeke_new_2012].
The problem arises in the balancing between interpretability and the higher
performance of the algorithms inspired by nature (such as neural networks). From
a business perspective, dropout prediction should also be considered as a
business objective, which requires more than predicting if the customer will
churn or not [@devriendt_why_2019], where higher interpretability provides
better support in the development of retention strategies. The developed SLR
also raises the possibility of integrating different algorithms using ensemble
methods or integrating several models using a hybrid approach. None of the
studies integrated the survival approach to predict customer dropout, for
example, using a hybrid approach.

It is considered positive if actions are developed to retain customers, but the
problems should also be considered, such as the following: (1) customers who
have greater risk of dropout should be targeted to provide a base for a better
ROI in the retention strategies [@xie_customer_2009; @coussement_churn_2008] and
(2) the retention strategies should be developed focused on customers with
higher satisfaction, or its inclusion could be a reminder of the contractual
agreement nearing an end and could lead to churn [@devriendt_why_2019].

From RQ3, several types of features being used were able to be identified, such
as demographic, behavioral, and economic indicators, pictorial data, network
relationships or high cardinality features. The problem that arises is that some
studies used data and features that were not described, and this creates a major
issue, How can reproducibility be developed in a study without the availability
of the data or the identification of the features used? Considering that science
is driven by data, with the development of new technologies, the increasing
complexity of research and the amount of data collected, the challenge is to
ensure that research is available to all [@Hanson_Sugden_Alberts_2011]; this
requires both availability of the data and the algorithms so that they can be
explored by other researchers. The features are selected mainly to verify the
performance of the models, and are essential to performance prediction,
accuracy, and the steps for processing the data, which are fundamental to
improve the model accuracy [@azeem_churn_2017].

There are several challenges around the timing related to dropout, or
considering the dynamic behavior of the customer in the intent to drop out
[@alboukaey_dynamic_2020]. The importance of understanding when dropout will
occur and the risk when discarding the temporal perspective of the problem seems
to be an element that should be addressed. Few studies considered this
[@perianez_churn_2016; @burez_separating_2008]. This shows an opportunity to
address the importance of the timeframe and its influence on the efficiency of
the model.

According to each business model, the timeframe could be addressed considering
the survival probability according to the customer relationship age, and dropout
predictions could be developed according to these survival probabilities, as
suggested by @esteves_churn_2016, to investigate which data timeframe produces
the best result and how the efficiency of the models is influenced by this
timeframe. Exploring the duration of the relation and the understanding of the
features that increase or decrease that duration seems to be an important
approach that could complement the existing approaches to predicting dropout.

From RQ5, the literature analysis showed that different types of questions
arise. Which are the best approaches to develop the analysis of the performance
in predicting dropout? Several metrics are customer dropout is to improve the
performance of organizations in retaining customers, which is a management
problem in which data mining is adopted [@verbeke_new_2012]. The goals of the
model should be formulated considering the context of the problem that is being
addressed; in marketing retention strategies, the up-lift supports the
development of proactive actions to minimize the investment in retention
strategies [@coussement_churn_2008]. Some assumptions that underlie the
adoption of uplift metrics consider that customers with a higher risk of
churning could not be the best targets, as suggested by @ascarza_retention_2018.
Other researchers addressed the problem using the top-decile lift to develop
more proactive actions to retain the customers at risk of churning
[@coussement_churn_2008;@xie_customer_2009]. This approach considers the 10% of
customers with more risk, and investments in retention strategies should be
developed that distinguish the churners susceptible to marketing actions from
those who will leave anyway [@coussement_comparative_2017]. Although uplift
models seem to be good strategies, they should also used, such as AUC,
sensitivity, specificity, recall, precision, and F-score. However, the goal
ofconsider factors other than risk and customer satisfaction, as not taking this
into consideration could be counterproductive and the model should be removed
from the retention strategy.

The true business objective is to reduce customer churn. Customers who are about
to churn but cannot be retained should be excluded from the campaign, as
targeting them will be a waste of scarce resources [@devriendt_why_2019]. Using
these models seem to be a good strategy, as they can outperform predictive
models that consider only accuracy from a profitability busshould be considered
that customers with a higher risk of churning may not be the best targets to
develop retention strategies. Those perspectives entail the dropout.

that a business context, or the clarification of a business objective underlying
the prediction of customer dropout, should be developed, to clarify which
objectives should be achieved before employing the profitability of reducing g
machine learning algorithms. Surprisingly, the analyzed studies did not address
the customer lifetime value as an objective to optimize considerininess
perspective.

# Aspects to consider

-   Interpretability from RQ2
-   The business objective is to increase the number of members and organization
    profits
-   piping several algorithms to improve accuracy. Aka hybrid approach
-   @Alboukaey_dynamic_2020 proposes ...
-   grep the articles addressing hybrid: pdfgrep -ri "hybrid.{1,10} approach"

# Results

In this section, we present our experiments to validate the proposed models,
comparing against other approaches. The models where optimized using the
hyper-parameters Grid Search technique. The explored hyper-parameters and the
best values of these parameters for every model are listed in (Table 3).

```{=latex}
\begin{table}[ht]
    \footnotesize
    \centering
    \begin{tabular}{p{0.15\textwidth}p{0.55\textwidth}p{0.25\textwidth}}
    \hline
    \textbf{Model name} & \textbf{Explored parameters values} & \textbf{Best parameters} \\
    \hline
    Survival trees  & pysurvival random forest                  & a \\
    Survival trees
    with clusters   & pysurvival random forest with clusters    & a \\
    Scikit survival 
    trees           & scikit survival                           & a \\
    Scikit survival 
    with clusters   & scikit with clusters                      & a \\
    Scikit survival 
    gradient boost  & scikit survival gradient boost            & a \\
    Scikit survival 
    gradient boost 
    with clusters   & scikit with clusters                      & a \\
    \hline
    \end{tabular}
    \caption{Hyper-parameters best values}
    \label{hyperparametersbestvalues}
\end{table}
```
The results of the performance of the models are available in table 4. Colocar o
modelo. Resultados do modelo

```{=latex}
\begin{table}[ht]
    \footnotesize
    \centering
    \begin{tabular}{p{0.15\textwidth}p{0.35\textwidth}r p{0.10\textwidth}}
    \hline
    \textbf{Model name} & \textbf{Results} & \textbf{n}\\
    \hline
    Survival trees  & RMSE 57.815                                     & 25316\\
                    & MAE 18.966                                      & \\   
                    & MEAE 38.557                                     & \\
    Survival trees  & Cluster 0: RMSE  2.785 MAE  0.000 MEAE  0.899   &   930\\
    with clusters   & Cluster 1: RMSE 78.141 MAE 28.537 MEAE 56.989   & 17067\\
                    & Cluster 2: RMSE 15.081 MAE  1.995 MEAE  6.850   &  2423\\
                    & Cluster 3: RMSE  9.354 MAE  2.506 MEAE  6.427   &  2816\\
                    & Cluster 4: RMSE  8.805 MAE  1.349 MEAE  3.725   &  2080\\

    Scikit survival 
    trees           & scikit survival                                 & \\
    Scikit survival 
    with clusters   & scikit with clusters                            & \\
    Scikit survival 
    gradient boost  & scikit survival gradient boost                  & \\
    Scikit survival 
    gradient boost 
    with clusters   & scikit with clusters                            & \\
    \hline
    \end{tabular}
    \caption{Hyper-parameters best values}
    \label{hyperparametersbestvalues2}
    * Root Mean Square Error (RMSE), Mean Absolute Error (MAE), Median Absolute Error (MEAE)  
\end{table}
```

# Conclusion

Article Ascarza

-   Retention Futility: Targeting High-Risk Customers Might be Ineffective
    [@ascarza_retention_2018]

Ascarza, E. (2018). Retention Futility: Targeting High-Risk Customers Might be
Ineffective. Journal of Marketing Research, 55(1), 80-98. sim.
<https://doi.org/10.1509/jmr.16.0163>

Example of Developed actions place in the discussion:

``` markdown
Each month, the company identified the customers who were up for renewal and
split them (randomly and evenly) between a treatment group that received a 
"thank you" gift with the letter and a control group that received only the 
renewal latter.
```
-->

# References {.unnumbered}

::: {#refs}
:::

# Appendix: Chunk options {.unnumbered}

## Software versioning R

### R

```{r fig-versioning, echo = TRUE}
cat(paste("#", capture.output(sessionInfo()), "\n", collapse  = ""))
  # or use message() instead of cat()
```